---
title: "CourseraPMLWk3"
author: "Issa Coulibaly"
date: ""
output: html_document
---

```{r,echo=FALSE}
path<- 'C:/Users/IICOUL/Documents/Coursera/Practical Machine Learning/Week_3/'
setwd(paste(path))
load(paste(path,"Coursera_ML_Week3Project.RData", sep=""))

```


```{r,eval=FALSE,echo=FALSE}

train <- read.csv("pml-training.csv", header=TRUE,stringsAsFactors=F, strip.white = TRUE,na.strings=c("","NA"))
test <-read.csv("pml-testing.csv", header=TRUE,stringsAsFactors=F, strip.white = TRUE,na.strings=c("","NA"))
dim(train);dim(test)
```

Taking a look at variable summary
The output is too long to be displayed here. Some variable appear to have many missing data. Others are character variable aside from the classe variable.
```{r,eval=FALSE}
summary(train)
```
Removing near-zero variables
```{r}
library(caret)
nzv<-nearZeroVar(train, saveMetrics=TRUE)
sum(nzv$nzv)
train2<-train[,!nzv$nzv]
test2<-test[,!nzv$nzv]
```


Removing variables with too many missing data
```{r}
n2<-dim(train2)[1]
missd2<-sapply(train2, function(x) sum(is.na(x))/n2)
mssta<- missd2 > .15
train2<-train2[,!mssta]
test2<-test2[,!mssta]
```

Checking predictors relationships visually.
Only a few examples are provided here.

```{r}
qplot(X,num_window,colour=classe,data=train2)
qplot(X,roll_belt,colour=classe,data=train2)

```
Removing time variables  and other variables.
```{r}
train2$raw_timestamp_part_1 <- train2$raw_timestamp_part_2 <- train2$cvtd_timestamp <- train2$user_name <- train2$X <- NULL
test2$raw_timestamp_part_1 <- test2$raw_timestamp_part_2 <- test2$cvtd_timestamp <- test2$user_name <- test2$X <- NULL

```

Removing highly correlated variables
```{r}
cor.m <-cor(train2[, -54])
cor.var<-findCorrelation(cor.m)
train2<-train2[-cor.var]
test2<-test2[-cor.var]
```


Training model using random forests approach
Predictor variables were centered and scaled 
```{r,eval=FALSE}
rfFit<- train(classe ~ ., method="rf",preProcess=c("center","scale"),
n.tree=500,data=train2)
```

Taking a look at the output 
```{r}
rfFit

```

The error rate is
```{r}
1 - max(rfFit$results$Accuracy)
```
Plot accuracy
```{r}
plot(rfFit)
```

Most important variables in the prediction model
```{r}
varImp(rfFit)
```

Predicting the 20 cases from the test set
```{r,eval=FALSE}
rfPred<-predict(rfFit,newdata=test2)
```
```{r}
rfPred
```

Writing out prediction files
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(rfPred)
```




